---
title: "Reproducible Research Week 2 / Project 1"
author: "Herna Osorio"
date: '2022-05-14'
output: html_document
---

# Reproducible Research Week 2
for the first part of the project I make sure that I load the appropriet data for the project.

```{r}
  dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
      download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
      unzip('repdata%2Fdata%2Factivity.zip')
```

After loading the data I make sure to read it:
```{r}
data <- read.csv("activity.csv")
```

Now i generate a historgram of total steps per day:
```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green",xlab="Number of Steps")
```

after that I go ahead and calculate the mean and the median:
```{r}
rmean <- mean(steps_by_day$steps)
rmean

rmedian <- median(steps_by_day$steps)
rmedian
```

after all these calculations we go ahead and start generate the daily activity pattern:

```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

```

no we Impute the missing values, this is by far the most complicates part of the analysis:
```{r}
NATotal <- sum(!complete.cases(data))
NATotal

StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
      obs <- data[i, ]
      if (is.na(obs$steps)) {
            steps <- subset(StepsAverage, interval == obs$interval)$steps
      } else {
            steps <- obs$steps
      }
      fillNA <- c(fillNA, steps)
}

new_activity <- data
new_activity$steps <- fillNA

StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Histogram to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)
```

now we look at the differentes steps per weekday or weekend:

```{r}
rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal

rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal

rmediandiff <- rmediantotal - rmedian
rmediandiff

rmeandiff <- rmeantotal - rmean
rmeandiff

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```

and this concludes the course project
